Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2017-03-08 Referee response.tex

Response to Referee Report on:"Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach", Sahar Rahmani et al. 
We thank the referee for her/his helpful comments and have addressed them individually below in lines beginning with "-". Changes to the text are highlighted in bold in the manuscript and also appear in quotations after our responses below. To improve clarity, we made some minor changes in the text which were not requested by the referee; these are in red bold highlighted text.

We hope changes made to the paper, makes it more clear why SOMs is useful method for classifying galaxy spectra and how it does not consider as an arbitrary classification.

"Most current classification methods cannot classify SEDs outside the range of templates, and will fail in special cases." (abstract)
Any classification method can generate a classification for any input.  It may be the case that this is of low confidence, or it is incorrect, but there is still some output.  I suggest rephrasing this claim to indicate the actual limitation (low accuracy? low confidence? something else?).
	- We rephrase the sentence as follows:
	    "Most current classification methods classify spectra outside the
		range of templates with low accuracy, and will fail in special cases"

In the end, how does one use the output of the SOM classification? Can you evaluate its accuracy in some way?  I understand that SOMs are exploratory and unsupervised, but there should be some demonstrable benefit or value from having done it.
    - There is no standard method to measure the accuracy of the SOM classification, however we hope that our description about comparing K-means clustering and self-organizing maps (see below and Section 4.3) sheds light on the usefulness of the SOM method. We find that, compared to K-means, SOM seems to be better at both discerning differences between similar templates and classifying objects which do not exactly match the templates.

The only statement I can see about something learned from the exercise is that the SOM yielded results with higher correlations between non-spectral properties.  But correlation is never quantified; instead plots are shown, leaving the reader's eye to estimate correlation.  To my eye, the plots in the lower row of Figure 10 have low correlation, if any.  What is the R^2 value?  And why does this matter? 
    - We added the following sentences to sections 4.1, 4.3 and 5 to elaborate on the usefulness of the SOM method:
    Section 4.1:
	    "One of the key features of SOMs is the ability to not only classify objects but to provide an {\em ordered} classification. 
       Figure~\ref{fig: props_vs_props} shows that the SOM correctly determines an ordering of the physical properties of the galaxy models used to make the classified spectra. 
       The median properties of the spectral groups change on going from one end of the SOM to the other.
       K-means clustering and supervised ANNs can sort galaxy spectra into classes but do not naturally order them."

	    Section 4.3: 
	    "One of the other key features of SOMs compared to supervised classification is the ability to both interpolate between and extrapolate outside of training set classes.
        This permits a fuller exploration of dataset properties and detection of outliers."

	    Section 5:
        "This particular sample of high-redshift galaxies is well-described by the range of the K96 templates; some of the spectra fall in between template classes.  This implies that these galaxies are not markedly different from those found in the local universe. The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relation.
     Other classification methods such as K-means and supervised artificial neural networks can provide a general classification with similar performance to the SOM. The ordered results and intermediate classes produced by self-organizing maps are distinct advantages of this method which can be utilized for other astronomical applications."

        - Regarding the correlation: in Fig. 9 the main message is how the colour of the points (corresponding to the classification) is ordered within each plot, rather than the actual correlation between parameters in the plot.

In response to major problems:
1.	Motivate the use of 1D versus 2D SOMs (or both).  As currently presented it is not clear what the benefit is of using either one, especially since they give different results.  Which one is better? Since they have different results, what does each one tell us?  At the end of section 4, the paper advises investigators to use the 2D approach first.  Why?  And if so, what additional knowledge is gained by applying the 1D approach? 
	- We add the following sentences to the end of section 4:"After finding and removing outliers in the input data, the use of 1D and 2D maps is complementary. One-dimensional maps are helpful for exploring the general behaviour of the data, while 2D maps show a more detailed picture."
	We also add the following sentences to section 4.3 to compare 1D and 2D maps:
    	"T12 showed that by masking emission lines (i.e., by ignoring details in spectra) two types of galaxies, Sb and SB6, can be misclassified as one another.
        As mentioned before, 1D SOMs are a good method to obtain a big picture of the problem under study and do not consider the small details. 
        Therefore, we can see that the SB6 category is placed near Sb galaxies in the 1D network in Fig.~\ref{fig: 1by22T}.
        However, in the upper right panel of Fig.~\ref{fig: 12by12}, we can see that the Sb and SB galaxies are completely separated. This shows that more details of spectral shapes are considered by 2D networks, which is one of the main advantages of  using 2D maps over 1D ones."

2. Although the paper explains why the SOM has more flexibility/expressivity than the neural network approach in reference T12, it does not compare the SOM to any other method out there.  The obvious candidate for comparison would be to apply k-means clustering (e.g., cluster the 142 spectra and then each clusters to its closest match of the original 12 templates).  But to compare the meaningfully, again you'll need some measure of performance, which currently is lacking.
    - We added Section 4.3 to the paper, which shows the result of K-means analysis on the data. 
    We classified 12 template galaxies in 4 groups and compared them with the classification from SOM. We also classified 142 galaxies using K-means clustering and matched them with the nearest K96 template using spectral angle distances. We compare the classification results with SOM classification and show their similarity and dissimilarity.

In response to minor problems:
The "mock sample" in section 3.3 does not add anything to the paper.  I suggest omitting it.  It raises more questions than it answers.  For example, is this randomly generated data?  Why 27 samples? Apparently there are only two features with only two possible values each, so there are only 4 unique possible samples, so 23 of the 27 must duplicate some other sample, and the data set is strongly over-specified.  Further, the features have values that are categorical rather than numeric (e.g., "high" and "low"), and it is not at all clear how the SOM, which employs Euclidean distance, can handle that kind of data.  Being able to divide a data set with only 4 unique samples into 4 distinct groups seems to be a non-accomplishment.
    - We removed this section.

I could not understand how the 2D SOM was constrained so that the results in the right-hand side of Figure 11 were obtained.  This network has the same topology as the one on the left yet somehow the K96 templates "are in a small region" (p. 10, line 43, column 2).
    - We used another SOM library to create the second type of SOM network. In the second method, by controlling the number of neighbourhoods, we can squeeze all the K96 templates into a small part of the map. In this case we can check for outliers in the sample galaxies, of which none were present here.

Euclidean distance is not the best way to match spectra; I recommend using spectral angle distance instead.
    - The main advantage of the spectral angle distance over the Euclidean distance for finding differences in spectra is that in the spectral angle distance looks at the general shape of the spectrum. Although SOM uses Euclidean distance to find the differences in some layer of classification, it also considers the overall shape of the spectrum. 
    Therefore, we decided to not use the spectral angle distance for SOM method.
    However, we used spectral angle distance for the K-means clustering and for finding the best match between K-means clusters and the templates in Section 4.2.

This isn't a fully unsupervised approach, since the SOM results are "classified" using the 12 template types (supervision).  That's not a weakness, but it means it would be more accurate to refer to the  method as "weakly supervised" or "partially supervised."
    - We add the following sentences to Section 1 where we introduced self-organizing maps: "An SOM can be trained without supervision and its results used to classify other data. Therefore, this method can be considered as (semi)-supervised."

Section 4.1.1 (p. 6, line 37-39, column 2) states that "The black color indicates that the left neuron is completely different from the other two groups" but earlier it was explained that these colors are relative to the range of differences present.  So black and white could be quite close (e.g., 0.1 and 0.2) or they could be large (e.g., 1 and 100000) but there's no way to know from the color.
    - The referee is correct: the differences are relative and not absolute. We change the phrase to:
        "They are more similar to each other than to the ones in the left neurons".

Section 4.1.1 (p. 7, lines 10-11, column 1) confusingly states that Figure 5 shows that "5 groups on the left-side neurons are separated from 7 groups on the right side of the map with two dark-grey colors between them."  There are indeed 5 groups then 2 dark-grey links then 7 groups, but those 7 are broken into sub groups with even greater contrast (black links).  So it would be more appropriate to say that the galaxies are split as 5 - 1 - 1 - 5 or even 6 - 1 - 6.
    - We changed the lines to the following:
        "Five groups in the left-side neurons are separated from the rest of the groups with two dark-grey colours between them. Five groups in the right-side neurons are separated from the rest by the black colour, and Sc and SB6 galaxies are in two separate blocks. This figure is a good example of the problem of using one-dimensional SOMs to explore the data. Since are the templates are forced to be arranged in a linear array, the SB6 and Sc groups are isolated and not connected to any other groups. However, in a two-dimensional map, due to the connection of each neuron to 6 more neurons, template galaxies have more degrees of freedom to find their neighbourhood.

@@

Section 4.1.2 (p. 7, line 9, column 2) says that Figure 8 has 12 plots, but actually it has 14.
    - Changed it to 14.

It is hard for the reader to track K96, T12, and N09 when their definitions are embedded pages before they are referenced.  I suggest including a small table with these labels, what they are, and their full citation so readers can quickly understand what is meant.
    - Unfortunately, we could not find any suitable place to put the suggested table. However, we made sure that all of the K96, T12 and N09 abbreviations are hyperlinked. In the PDF version, the reader can click on them to see the actual citation.

When re-using figures from a previous publication, you may need copyright permission from the journal source (e.g., Figure 1 was published by the Astronomical Journal).
    - Thank you for reminding us about the copyright permission. We checked the policy of the Astronomical Journal and for copyright purposes the permission of the author is enough. The sole author of the T12 paper is a co-author to the current paper and we have his permission to use the figure.

### 2017-03-08 answer_to_second_referee_comments.tex

Reviewer's Comments:

This paper proposes the use of self-organizing maps (1D and 2D) to aid in classifying galaxy spectra.  The revision has addressed several points and improve the clarity of the text.  However, there are several remaining issues.

1. The authors did not address my comment about the significance of the work.  What was learned?  What method is advocated, and what is the evidence for its superiority?

The sentence that was added about 1D vs. 2D maps comes very late in the paper and is quite abstract, stating that the two methods are complementary.  The Introduction should establish the goals of the investigation so the reader understands the purpose of each experiment as well as its significance.
%20170122PB: draft added
-- We have reorganized the last two paragraphs of Section 1 to better describe the goals of the investigation.

The paper claims that "Compared to previous attempts to classify this sample, the self-organizing map approach was able to better classify the spectra which had similarities with more than one template." (abstract), but I cannot find a quantitative comparison with any previous results.  Figure 9 is presented as the result to compare with previous work, but since the previous results are not presented nor summarized, the reader can't tell how they differ.  The paper reports "more precision", but this is not defined; does it mean smaller error bars?  Stronger correlation?  How does this show "better classification"?  To support such a claim, this paper needs to select a performance metric and report on the values achieved in these experiments, compared to other methods.
--- should I put the figure from the previous paper, maybe Noll09 one??! 
% I don't think including the figure from Noll09 will be enough. It's the performance metric that's critical here.

I had suggested comparing the results to a simple nearest-neighbor classification baseline, but this was not addressed.
%--- why could not I find anything about it in the 1st report?:(
%20170122PB: it's there but well-hidden - the sentence which includes "what performance does one get simply by assigning each of the 142 spectra to its closest match from the 12 templates"

The comment about how SOMs provide an ordered clustering of the data (with respect to feature values), which k-means does not, is certainly true.  But the paper must then explain why this is important or valuable.
--- % need comment here

2. Perhaps I have missed it somewhere, but I don't think the paper ever states how exactly the classification of new data is done.  As I understand it, the authors train an SOM using template spectra, then classify new spectra by determining which SOM node is the best fit for the new item and classifying it based on the labeled template spectra that also reside in that node -- but this isn't actually stated anywhere.  Please add text explaining the classification step.  This would probably go in the final paragraph of the introduction and/or a new subsection in section 3 ("Method") - this section describes SOMs and k-means but not how to use either model to classify new data. (i.e., the SOM can't classify anything without knowing the labels of its original data set.  So although it builds the SOM in an unsupervised way, it requires the supervision of the template labels to be able to assign a label to a new item).
-- Added the following paragraph  at the end of section 3.2: %20170122PB: edited for grammar, inserted in paper text. Needs a bit more discussion.
    "As mentioned in the introduction, one of the main advantages of the SOM method  is that the resulting networks can be used to cluster new datasets with no additional training required. 
     New data with the same dimensionality as the original input data can be presented to the already trained network: the SOM algorithm finds the best matching unit considering the weight of nodes in the trained network and each vector from the new data set.
     The winner node determines the place of each new vector on the SOM.
     Based on the location of the new vectors on the SOM, they can be compared with the original data set. %20170122PB: unclear what "they" refers to
     Sections~\ref{sec: 1Dv} and ~\ref{sec: 2D} show the results of galaxy spectral classifications using networks trained on template spectra."

On page 2, lines 29-31, the authors added this text: "An SOM can be trained without supervision and its results used to classify other data. Therefore, this method can be considered as (semi)-supervised." As written, the phrase "this method" seems to refer to SOMs (so the statement wouldn't be true - SOMs are fully unsupervised), but I think instead the authors' proposed method (of building an SOM and then applying it to classify new data, using labels on the SOM-grouped data) is what is meant.  Please clarify.
-- Changed to: %20170122PB: edited for grammar, inserted in paper text
    "The training of a self-organizing map is fully unsupervised.
     Using the resulting map as a template to classify other data requires labelling the output, which is why some groups consider self-organizing maps to be a (semi)-supervised method."

3. The inclusion of k-means helps give useful context to this work. Figure 10 shows a head-to-head comparison of k-means and SOM applied to the 12 templates using 4 clusters/nodes.  But I can't tell which one is better.  The text states that they are similar but "show discrepancies for spectra types between the extreme cases".  Which ones?  How can I tell from this figure?
-- % need a comment here

At the bottom of page 10, the paper states: "The other difference between these two methods is that the network created by the SOM method can be used to classify the other spectra, while for the K-means clusters they can only be used for one set of data."  This isn't true.  (The paper itself points out that SOM reduces to k-means if the weight on the neighborhood function goes to 0, so you should be able to use the result in the same way.)  You can (and should, to do a direct comparison) use the clusters just like you used the SOM nodes - for a new data point, assign it to the most similar cluster center, then assign the template label from the original data used to construct the clusters.
--- I do not know how to do that!
% for each spectrum to be classified, compute the distance (in whatever metric we end up using)
% between the spectrum and the K cluster centres. Spectrum is classified as belonging to whatever cluster results in the smallest distance.

To analyze the 142 galaxies, the SOM and k-means experiments used different methodology and so are difficult to compare or understand. It seems that for k-means, the authors clustered the 142 sample galaxies and then classified the mean spectrum in each cluster and reported how well that did.  But for the SOM, the authors constructed the SOM based on the 12 templates from K96 and then assigned classes to individual new spectra (not the nodes).  This doesn't seem comparable.  Please use the same methodology in both cases.
--- again I do not know how to used the clusters classified based on template to classify new data sets.
% this is a fair comment - need to think more about how to compare

Further, k-means and the SOM used different distance functions (Euclidean vs. spectral angle distance) when operating on the same data.  Differences in results could simply be due to this change.  The authors' explanation for this choice does not make sense to me.
--- We have results from the Euclidean distance for the K-means! But I steel think because of the way SOM operates it is better to use SAD! Or maybe describe it better!

The text says that Figure 11 shows "the best and the worst fitting results among the 22 clusters" (for k-means) but instead it seems to show the best-fit for two clusters (9 and 22) which both look pretty good.  The text implies that Fig 11 (right) is a poor match, but it looks very good.  Please explain.
--- We apologize: this was a typo with "right" and "left" switched.

%%%%! I wrote right and left wrong again:( I should be extra careful on this one.

4. If k-means is retained, the way it is presented needs improvement. The description of k-means clustering in section 1 is a nice addition, but the comparison between k-means and SOM is confusing and wordy. K-means optimizes a partition of the data based on distance in the feature space only.  SOMs operate on a combination of distance in feature space and distance in the resulting low-dimensional map.  I am not sure how this achieves "smoothing", nor whether this is an important detail (the paper never mentions it again).  I recommend explaining what is meant here and why a "smooth" solution is better, or omitting this detail.  Explaining how SOM can be reduced to k-means is a good addition.
% PB: I edited this a bit, but not quite done as of 20170123 - need to talk about the smoothing part.

Please a add citation to the original k-means algorithm (e.g., MacQueen, 1967) when it is first mentioned (in addition to citing people who've used the algorithm).
-- Added the proper citation.

Section 3.3 outlines the k-means method that is used in the experiments.  The language here appears to be more informal and less polished than that which describes the SOM method.  To explain k-means with the same level of detail, it would be appropriate to show the objective function that is employed and to explain that it uses an alternating optimization to assign points to centroids and then update the centroids, iterating until... what?  (There are different termination conditions used by different implementations, so it would be useful to indicate here what the matlab method does.)

--- After attending to the #3 %(PB: I'll help with the language part)

### 2017-03-08 main.tex
change intro part on how using K -means and what is the different between K-means and SOM

\newcommand \aaj {A\&A}
\newcommand \aarv {A\&ARv}%: Astronomy and Astrophysics Review (the)
\newcommand \aas{A\&AS}%: Astronomy and Astrophysics Supplement Series
\newcommand \afz {Afz}%: Astrofizika
\newcommand \aj {AJ}%: Astronomical Journal (the)
\newcommand \apss {Ap\&SS}%: Astrophysics and Space Science
\newcommand \apj {ApJ}
\newcommand \apjs {ApJS}%: Astrophysical Journal Supplement Series (the)
\newcommand \araa {ARA\&A} %: Annual Review of Astronomy and Astrophysics
\newcommand \asp {ASP Conf. Ser.}%: Astronomy Society of the Pacific Conference Series
\newcommand \azh {Azh}%: Astronomicheskij Zhurnal
\newcommand \baas {BAAS}%: Bulletin of the American Astronomical Society
\newcommand \mem {Mem. RAS}%: Memoirs of the Royal Astronomical Society
\newcommand \mnassa {MNASSA}%: Monthly Notes of the Astronomical Society of Southern Africa
\newcommand \mnras {MNRAS} %: Monthly Notices of the Royal Astronomical Society
%\newcommand {Nature}%(do not abbreviate)
\newcommand \pasj {PASJ}%: Publications of the Astronomical Society of Japan
\newcommand \pasp {PASP}%: Publications of the Astronomical Society of the Pacific
\newcommand \qjras {QJRAS}%: Quarterly Journal of the Royal Astronomical Society
\newcommand \mex {Rev. Mex. Astron. Astrofis.}%: Revista Mexicana de Astronomia y Astrofisica
%\newcommand {Science }%}%(do not abbreviate)
\newcommand \sva {SvA}%: Soviet Astronomy
\newcommand \aap {APP} %:American Academy of Pediatrics
\newcommand \apjl {ApJL} %:The Astrophysical Journal Letters

@@

\begin{abstract}
The \textbf{\textcolor{red}{spectrum}} of a galaxy contains information about its physical properties.
Classifying \textbf{\textcolor{red}{spectra}} using templates helps elucidate the nature of a galaxy's energy sources.
However, available templates are limited, which can complicate \textbf{\textcolor{red}{spectral}} classification.
Most current classification methods classify \textbf{\textcolor{red}{spectra}} outside the range of templates \textbf{with low accuracy}, and will fail in special cases.
In this paper, we introduce a new method of classifying the \textbf{\textcolor{red}{spectra}}
of high redshift galaxies: the self-organizing map.
Self-organizing maps are a type of artificial neural network, mostly used to identify subgroups in a sample.
We trained semi-supervised self-organizing map networks using a set of
templates covering the wavelength range from far ultraviolet to near infrared. 
The trained networks were used to classify the \textbf{\textcolor{red}{spectra}} of a sample of 142 galaxies with $0.5 < z < 1$.
Compared to previous attempts to classify this sample, the
self-organizing map approach was able to better classify the \textbf{\textcolor{red}{spectra}} which had
similarities with more than one template.
The \textbf{\textcolor{red}{median spectra}} and physical properties of the galaxies in each neuron of the self-organizing maps are consistent with other results on galaxy evolution.

\end{abstract}
\begin{keywords} 
 galaxies: high-redshift, 
 galaxies: \textbf{\textcolor{red}{spectra}}, 
 methods: observational, 
 methods: statistical, 
 methods:data analysis
\end{keywords}

@@

\import{sections0.07/}{intro0.07.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%DATA
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\import{sections0.07/}{data0.07.tex}

@@

\import{sections0.07/}{method0.07.tex}

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Results
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\import{sections0.07/}{results0.07.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Summery
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Summary and Future Applications}
\label{sec: summary_SOMZ}

Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.)
Galaxy spectra can be classified in various networks with different sizes/dimensions based on the information needed from the data. 
If a broad and general classification is required, networks can have one dimension with a few neurons. 
\textbf{\textcolor{red}{Other classification methods such as K-means and supervised artificial neural networks can provide a general classifications with similar performance as SOM.}}
If one needs more detailed classifications, a higher number of neurons should be used.
Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
%Some small differences could easily be the result of atmosphere fluctuation or other instrumental or observational problems. 
When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

@@

A sample of 142 high-redshift galaxy \textbf{\textcolor{red}{spectra}} from \citetalias{Hossein12} was classified by the trained networks.
The classification results showed that using SOMs can allow identification of galaxies with \textbf{\textcolor{red}{spectra}} similar to two or more morphological types.
The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks.
\citetalias{Hossein12} used a supervised training method to train networks with \citetalias{Kinney96} \textbf{\textcolor{red}{spectral}} templates and classify the same sample of 142 galaxy spectra;
however, \textbf{\textcolor{red}{this method}} could not classify 37 out of 142 spectra.
The unsupervised artificial neural network method used in this project was able to classify  all 142 spectra in the sample
as belonging to one of the morphological types introduced by \citetalias{Kinney96} or a class in between those types.
\textbf{Although, it might seem that the classification spectra using SOMs are an arbitrary classification, we should remind that this classification is emerged form the results of data mining on the previous templates.}
%The classification based on SOM networks are not arbitrary, because there is an actual relation between each neurons}

The properties of the galaxies in the \citetalias{Hossein12} sample using the new classification were found to be more tightly correlated in mean values of age, specific star formation rate, stellar mass, and far-UV extinction than in previous studies. 
The properties of the galaxies in each group are in good agreement with their morphological types.
\textbf{This particular sample of high-redshift galaxies is well-described by the range of the K96 templates, although some of the spectra fall in between template classes. 
This implies that these galaxies are not hugely different from those found in the local universe.
The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relation.}

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for insightful comments which helped us to improve the manuscript.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada. 
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{apalike}
\bibliography{ref_mining_h.bib}

%\import{sections0.0/}{apps0.0.tex}
\import{sections0.07/}{apps_0.07.tex}
\end{document}

>>>>

\newcommand \aaj {A\&A}
\newcommand \aarv {A\&ARv}%: Astronomy and Astrophysics Review (the)
\newcommand \aas{A\&AS}%: Astronomy and Astrophysics Supplement Series
\newcommand \afz {Afz}%: Astrofizika
\newcommand \aj {AJ}%: Astronomical Journal (the)
\newcommand \apss {Ap\&SS}%: Astrophysics and Space Science
\newcommand \apj {ApJ}
\newcommand \apjs {ApJS}%: Astrophysical Journal Supplement Series (the)
\newcommand \araa {ARA\&A} %: Annual Review of Astronomy and Astrophysics
\newcommand \asp {ASP Conf. Ser.}%: Astronomy Society of the Pacific Conference Series
\newcommand \azh {Azh}%: Astronomicheskij Zhurnal
\newcommand \baas {BAAS}%: Bulletin of the American Astronomical Society
\newcommand \mem {Mem. RAS}%: Memoirs of the Royal Astronomical Society
\newcommand \mnassa {MNASSA}%: Monthly Notes of the Astronomical Society of Southern Africa
\newcommand \mnras {MNRAS} %: Monthly Notices of the Royal Astronomical Society
%\newcommand {Nature}%(do not abbreviate)
\newcommand \pasj {PASJ}%: Publications of the Astronomical Society of Japan
\newcommand \pasp {PASP}%: Publications of the Astronomical Society of the Pacific
\newcommand \qjras {QJRAS}%: Quarterly Journal of the Royal Astronomical Society
\newcommand \mex {Rev. Mex. Astron. Astrofis.}%: Revista Mexicana de Astronomia y Astrofisica
%\newcommand {Science }%}%(do not abbreviate)
\newcommand \sva {SvA}%: Soviet Astronomy
\newcommand \aap {A\&A} %:Astronomy & Astrophysics
\newcommand \apjl {ApJL} %:The Astrophysical Journal Letters

@@

\begin{abstract}
    The spectrum of a galaxy contains information about its physical properties.
    Classifying spectra using templates helps elucidate the nature of a galaxy's energy sources.
    However, available templates are limited, which can complicate spectral classification.
    Most current classification methods classify spectra outside the range of templates with low accuracy, and will fail in special cases.
    In this paper, we introduce a new method of classifying the spectra
    of high redshift galaxies: the self-organizing map.
    Self-organizing maps are a type of artificial neural network, mostly used to identify subgroups in a sample.
    We trained semi-supervised self-organizing map networks using a set of
    templates covering the wavelength range from far ultraviolet to near infrared. 
    The trained networks were used to classify the spectra of a sample of 142 galaxies with $0.5 < z < 1$.
    Compared to previous attempts to classify this sample, the
    self-organizing map approach was able to better classify the spectra which had similarities with more than one template.
    The median spectra and physical properties of the galaxies in each neuron of the self-organizing maps are consistent with other results on galaxy evolution.

\end{abstract}
\begin{keywords} 
 galaxies: high-redshift, 
 galaxies: spectra, 
 methods: observational, 
 methods: statistical, 
 methods:data analysis
\end{keywords}

@@

\import{sections0.08/}{intro0.08.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%DATA
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\import{sections0.08/}{data0.08.tex}

@@

\import{sections0.08/}{method0.08.tex}

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Results
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\import{sections0.08/}{results0.08.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Summery
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Summary and Future Applications}
\label{sec: summary_SOMZ}

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.)
    Galaxy spectra can be classified in various networks with different sizes/dimensions based on the information needed from the data. 
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

@@

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
    The classification results showed that using SOMs can allow identification of galaxies with spectra similar to two or more morphological types.
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks.
    \citetalias{Hossein12} used a supervised training method to train networks with \citetalias{Kinney96} spectral templates and classify the same sample of 142 galaxy spectra;
    however, this method could not classify 37 out of 142 spectra.
    The unsupervised artificial neural network method used in this project was able to classify  all 142 spectra in the sample
    as belonging to one of the morphological types introduced by \citetalias{Kinney96} or a class in between those types.
    \textbf{Although it might seem that the classes of spectra using SOMs are arbitrary, we remind the reader that this classification emerges from the results of data mining on the previous templates.}

    The properties of the galaxies in the \citetalias{Hossein12} sample using the new classification were found to be more tightly correlated in mean values of age, specific star formation rate, stellar mass, and far-UV extinction than in previous studies. 
    The properties of the galaxies in each group are in good agreement with their morphological types.
    This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates; some of the spectra fall in between template classes. 
    This implies that these galaxies are not markedly different from those found in the local universe.
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relation.
     Other classification methods such as K-means and supervised artificial neural networks can provide a general classification with similar performance to the SOM. The ordered results and intermediate classes produced by self-organizing maps are distinct advantages of this method which can be utilized for other astronomical applications.

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for insightful comments which helped us to improve the manuscript.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher, R.G. Abraham, M. Daley and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada. 
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{apalike}
\bibliography{ref_mining_h.bib}

%\import{sections0.0/}{apps0.0.tex}
\import{sections0.08/}{apps_0.08.tex}
\end{document}

### 2017-03-08 sections0.07/apps_0.07.tex
change intro part on how using K -means and what is the different between K-means and SOM

\newpage
\appendix
\section{Results of the 1D training and validation}
\label{app: 1d}
    \begin{figure}
        \begin{subfigure}[b]{0.5\textwidth}
            \centering
            \includegraphics[width=\textwidth]{../images0.01/1d/apps/dist_1_by_4.png}
            %\caption{$1\times4$ weight map}
             %\label{fig: 1by4T}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.5\textwidth}
             \includegraphics[width=\textwidth]{../images0.01/1d/apps/hit_t_1_by_4.png}
             %\caption{$1\times4$ hits map}
             %\label{fig: 1by4Thits}
        \end{subfigure}
                \caption{Results of training network in $1\times4$~grid.}
         \label{fig: 1by4T}
    \end{figure}

>>>>

%appendix
\newpage
\appendix
\section{Results of training for one-dimensional self-organizing maps}
\label{app: high_Z_1d_soms}
As mentioned in Sec.~\ref{sec: 1D_somz}, we changed the size of the SOM from $1\times2$ to $1\times22$. In that section we show some example of the results; here we show the rest of the maps to monitor the changes in the map over various sizes.

\label{app: 1d}
    \begin{figure}
        \begin{subfigure}[b]{0.5\textwidth}
            \centering
            \includegraphics[width=\textwidth]{../images0.01/1d/apps/dist_1_by_4.png}
            %\caption{$1\times4$ weight map}
             %\label{fig: 1by4T}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.5\textwidth}
             \includegraphics[width=\textwidth]{../images0.01/1d/apps/hit_t_1_by_4.png}
             %\caption{$1\times4$ hits map}
             %\label{fig: 1by4Thits}
        \end{subfigure}
                \caption{Results of training network in $1\times4$~grid.}
         \label{fig: 1by4T}
    \end{figure}
